---
title: "Sentiment Analysis with OpenAI Embeddings and R"
format: html
execute:
  echo: true
---

We begin by installing all necessary libraries. We also retrieve the environment variable OPENAI_API_KEY, which is set in a separate initialization script.

```{r setup}
#install.packages("glmnet")
library(tidyverse)
library(tidytext)
library(textstem)
library(caret)
library(glmnet)
library(jsonlite)
library(httr)

# --------------------------------------------
#  Set your API key - done outside of this code
# --------------------------------------------
openai_api_key <- Sys.getenv("OPENAI_API_KEY")
if (openai_api_key == "") {
  stop("Please set OPENAI_API_KEY environment variable.")
}
```

Now we read in our CSV datafile of 219.254 tweets regarding ChatGPT.

```{r}
data <- read_csv("file.csv")

# Expecting columns: recid, tweet_txt, label_txt

head(data)

```

The text of the tweets must be cleaned before sending to ChatGPT: removing non-letter characters and extra spaces, removing stop words, and lemmatizing.

```{r}
stop_words <- tidytext::stop_words

clean_data <- data |>
    mutate(text_clean = tweet_txt |>
    str_to_lower() |>
    str_replace_all("[^a-z]", " ") |>
    str_squish()) |>
    unnest_tokens(word, text_clean) |>
    anti_join(stop_words, by = "word") |>
    mutate(word = lemmatize_words(word)) |>
    group_by(recid, label_txt) |>
    summarize(text_clean = paste(word, collapse = " "), .groups = "drop")

head(clean_data)
```

We now define a function for retrieving the embedding of tweet content from ChatGPT via the OpenAI API.

```{r get-embeddings}
get_embedding <- function(texts, model = "text-embedding-3-small") {

    url <- "https://api.openai.com/v1/embeddings"

    response <- POST(
        url,
        add_headers(
            Authorization = paste("Bearer", openai_api_key),
            `Content-Type` = "application/json"
            ),
            body = list(
            model = model,
            input = texts
            ),
        encode = "json"
)

result <- content(response, as = "parsed", simplifyVector = TRUE)

embeddings <- result$data$embedding
return(embeddings)
}

```

The above function is called iteratively to build a data frame of embeddings.

```{r compile-embeddings}
batch_size <- 50
texts <- clean_data$text_clean
n <- length(texts)

emb_list <- list()

for (i in seq(1, n, by = batch_size)) {
    batch <- texts[i:min(i + batch_size - 1, n)]
    emb <- get_embedding(batch)
    emb_list <- append(emb_list, emb)
}

emb_matrix <- do.call(rbind, emb_list)

# Combine with labels

sentiment <- clean_data$label_txt
emb_df <- as.data.frame(emb_matrix)
emb_df$sentiment <- sentiment

```

It was expected that building the embeddings data frame would take a while. It took somewhere between 30 and 45 minutes.

We split the data 80/20 into training and test sets.

```{r split-data}
set.seed(123)
train_index <- createDataPartition(emb_df$sentiment, p = 0.8, list = FALSE)

train <- emb_df[train_index, ]
test  <- emb_df[-train_index, ]

```

We now do resampling with 10-fold cross-validation to estimate prediction of sentiment from embedded tweet content.

```{r training}
ctrl <- trainControl(method = "cv", number = 10)

model <- train(
    sentiment ~ .,
    data = train,
    method = "glmnet",
    trControl = ctrl
    )

model

```

On a laptop CPU, training took several hours. I was not expecting it to take so long so I did not build in a timer. This step probably should be farmed out to a GPU in the cloud in a production environment.

Model training weights are now applied to the test data to evaluate performance in a confusion matrix.

```{r confusion-matrix}
pred <- predict(model, newdata = test)
# confusionMatrix expects factor data when variables are strings
test$sentiment_fct <- as.factor(test$sentiment)
confusionMatrix(pred, test$sentiment_fct)
```

Accuracy of the model is 67%, which is significantly and substantially better than the naive classifier's 49% accuracy. Thus, using the derived model weights is a good, but admittedly not great, method for classifying a tweet's embeddings as having positive or negative sentiment. It would be interesting to see how model accuracy could be improved.